
# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

### Unzip if necessary
```{r}

if (!file.exists("activity.csv"))
{
unzip("activity.zip")
}
```
### Load the data
```{r}
activity<-read.csv("activity.csv", head=TRUE)
head(activity)
```
### Little bit of prepping
```{r}
activity.prepped<-activity
activity.prepped$date<-as.Date(activity.prepped$date)
activity.prepped<-subset(activity.prepped, !is.na(activity.prepped$steps))
head(activity.prepped)
```
### Sum steps by day
```{r}
activity.prepped.sum.by.date <- aggregate(list(steps = activity.prepped$steps), 
                                by=list(date=activity.prepped$date), FUN=sum)
head(activity.prepped.sum.by.date)
```

## What is mean total number of steps taken per day?

### Create histogram
```{r}
hist(x=activity.prepped.sum.by.date$steps ,col="red",breaks=25,xlab="Daily steps",
     ylab="Frequency",
     main="The distribution of daily steps")
```
### Calculate mean and median
```{r}
mean_steps<-as.integer(mean(activity.prepped.sum.by.date$steps))

median_steps<-as.integer(median(activity.prepped.sum.by.date$steps))
```
Mean = `r mean_steps` 
Median = `r median_steps` 

## What is the average daily activity pattern?
### Mean steps by interval
```{r}
activity.prepped.sum.by.interval <- aggregate(list(steps = activity.prepped$steps),   by=list(interval=activity.prepped$interval), FUN=mean)
head(activity.prepped.sum.by.interval)
```
### Create the plot 
```{r}
plot(activity.prepped.sum.by.interval$interval, 
     activity.prepped.sum.by.interval$steps, type="l",
     xlab="5' intervals", ylab="Average steps")
```
### Which interval has max number of steps
```{r}
max_steps<-max(activity.prepped.sum.by.interval$steps)
max_interval <- activity.prepped.sum.by.interval[activity.prepped.sum.by.interval$steps==max_steps,]
max_interval
```
## Imputing missing values
```{r}
Number_of_NAs <- nrow(subset(activity, is.na(activity$steps)))
Number_of_NAs
activity.impute <- activity
activity.impute.na <- is.na(activity.impute$steps)
activity.impute.sum.by.interval <- tapply(activity.prepped$steps, activity.prepped$interval, mean, na.rm=TRUE, simplify=T)
activity.impute$steps[activity.impute.na]<-activity.impute.sum.by.interval[as.character(activity.impute$interval[activity.impute.na])]
```
### Create histogram
```{r}
activity.impute.sum.by.date <- aggregate(list(steps = activity.impute$steps), by=list(date=activity.impute$date), FUN=sum)

hist(x=activity.impute.sum.by.date$steps ,col="red",breaks=25,xlab="Daily steps",
     ylab="Frequency",
     main="The distribution of daily steps")
```
### Calculate mean and median
```{r}
new_mean_steps<-as.integer(mean(activity.impute.sum.by.date$steps))
new_median_steps<-as.integer(median(activity.impute.sum.by.date$steps))
```
Mean = `r new_mean_steps` 
Median = `r new_median_steps` 

So mean is unchanged, median has small change.  
Looks like imputing causes higher frequency counts in the middle.


## Are there differences in activity patterns between weekdays and weekends?

### Add variable day and weekendindicator 
```{r}
Sys.setlocale("LC_TIME", "English")
activity.impute$weekday<-weekdays(as.Date(activity.impute$date))
activity.impute$weekendind<-ifelse(activity.impute$weekday=="Saturday" |activity.impute$weekday=="Sunday", "Weekend", "Weekday") 
```
### Mean steps by weekendind/interval
```{r}
activity.impute.wk <- aggregate(steps ~ weekendind+interval, data=activity.impute, FUN=mean)
head(activity.impute.wk)
```

### Create the plot 
```{r}
library(lattice)
xyplot(steps ~ interval | factor(weekendind),
       layout = c(1, 2),
       xlab="Interval",
       ylab="#steps",
       type="l",
       lty=1,
       data=activity.impute.wk)
```
  
Seems like people tend to wake up earlier weekdays but move more in the weekend.


